Query Optimization for Dynamic Imputation

### 2017-05-11 rebuttal.tex

\item W3. More experiments needed.

\item D1. Section 3, which is the meat of the paper, is very overloaded, without getting in depth in various parts of the system. It definitely should be split in more sections, and more details should be given. In particular:
-- Give some more information about cardinality estimation. Also, Algorithm 3 is too far from the text, and needs some explanation too.
-- Some more details are also required in the Imputation quality section. Why are you choosing these specific formulas for Loss, especially for the impute operator. Some examples would help.
-- Algorithms 1 and 2 could probably be presented in an easier to grasp way.

\item D2. It would be nice to show how exactly your approach can be implemented in an existing optimizer. This comes up in various places in the paper:
-- When defining the search space, why not first define the transformation operators that inject the imputation operators. Then any optimizer can pick these rules up and add them in its plan enumeration. Of course, the used approach that follows System-R's plan enumeration is useful, but why lose generality when presenting the search space. Moreover, showing the transformation rules can help you show the size of the search space that will better justify the restrictions you apply.
-- Why use simplified cost formulas for the traditional relational operators and not existing systems' formulas?
-- It would be great (but not necessary of course) to add your techniques in an existing DBMS, such as Postgres to see the impact on a real system in your experiments.

\resp{
Existing systems have very complex formulas for estimating query costs. We felt that the overhead of reimplementing an existing cost model would not be worth the benefit.
However, we added a new section that discusses in detail the main components for \ProjectName{}, so that developers of existing systems can use this as a roadmap
for implementation.

\todo{More sophisticated cost formulas from exiting systems could further improve our time estimates for traditional operators. In this case, we focused on developing the heuristics
relevant for the new operators, at the expense of simplifying implementation of common formulas, as we felt that was core to our contribution.}
}

@@

\item D3. Is there any way for the user to know what error to expect based on the $\alpha$ value? Moreover, is its value somehow proportional to the error (e.g., $\alpha=0.5$ would give roughly double the error of $\alpha=0.25$)? Or should the user simply try multiple execution with different values of $\alpha$ until the desired results are given?

  \resp{In our intended usage, the user starts by running with $\alpha = 0$ and increases $\alpha$ if the query takes too long to complete.
  \todo{
  We agree that having a way of providing an estimate on error would be interesting. Since error in this case relates to the imputation values
  generated by the whole-table imputation (as there is no ground truth), this would likely need to be on a case-by-case basis for distinct
  instantiations of the impute operator. 
  }
  }

### 2017-05-11 paper/algorithm.tex
algorithms pass, nitpicks

\section{Algorithm}
In order to correctly plan around missing values, we first extend the set of relational algebra operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) 
with two new operators (\Cref{sec:operators}). We then define the search space (\Cref{sec:search-space}) of plans to encompass non-trivial operations, such as joins and 
aggregations. Soundness of query execution is provided by our main design invariant, which guarantees traditional relational
algebra operators must never observe a missing value in any attribute that they operate on directly (\Cref{sec:placement}). 
The decision to place imputation operators is driven by our cost model (\Cref{sec:cost-model}), which characterizes a plan based on estimates for the 
quality of the imputations performed and the runtime performance. Our imputation planning algorithm is
agnostic to the type of imputation used (\Cref{sec:imputation}). Finally, we show
that while our algorithm is exponential in the number of joins (\Cref{sec:complexity}), in practice, planning times are not a concern.

@@

\subsection{Cost model}
\label{sec:cost-model}
The cost of a plan $q$ is a tuple $\langle \textsc{Penalty}(q), \textsc{Time}(q) \rangle$.
$\textsc{Penalty}(q) \in [0,1]$ is a heuristic that attempts to quantify the amount of information lost by the imputation procedures used.
$\textsc{Time}(q) \in (0, \infty)$ is an estimate of the runtime of a query $q$ derived from table statistics and selectivity estimation of the query predicates.

@@

\subsubsection{Cardinality estimation}
\label{sec:cardinal}
The computation of $\textsc{Penalty}(q)$ and$\textsc{Time}(q)$ relies on accurate cardinality estimates for the plan $q$ and its sub-plans.
These cardinality estimates are impacted not just by filtering or joining, as in the traditional relational algebra, but also by the imputation operators.
For example, a drop operator will reduce the cardinality of the result while an impute operator will maintain the same cardinality as the input.

@@

\begin{algorithm}
\input{algorithms/histogram-transform.tex}
\end{algorithm}

>>>>

\section{Algorithm}
In order to correctly plan around missing values, we first extend the set of relational algebra operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) 
with two new operators (\Cref{sec:operators}). We then define the search space (\Cref{sec:search-space}) of plans to encompass non-trivial operations, such as joins and 
aggregations. Soundness of query execution is provided by our main design invariant, which guarantees traditional relational
algebra operators must never observe a missing value in any attribute that they operate on directly (\Cref{sec:placement}). 
The decision to place imputation operators is driven by our cost model (\Cref{sec:cost-model}), which characterizes a plan based on estimates for the 
quality of the imputations performed and the runtime performance. Our imputation planning algorithm is
agnostic to the type of imputation used (\Cref{sec:imputation}). Finally, we show
that while our algorithm is exponential in the number of joins (\Cref{sec:complexity}), in practice, planning times are not a concern for most queries.

@@

\subsection{Cost model}
\label{sec:cost-model}
The cost of a plan $q$ is a tuple $\langle \textsc{Penalty}(q), \textsc{Time}(q) \rangle$.
$\textsc{Penalty}(q) \in (0,1]$ is a heuristic that attempts to quantify the amount of information lost by the imputation procedure used.
$\textsc{Time}(q) \in (0, \infty)$ is an estimate of the runtime of a query $q$ derived from table statistics, selectivity estimation of the query predicates, and the complexity of the
impute operation.

@@

\subsubsection{Cardinality estimation}
\label{sec:cardinal}
The computation of $\textsc{Penalty}(q)$ and$\textsc{Time}(q)$ relies on accurate cardinality estimates for the plan $q$ and its sub-plans.
These estimates are impacted not just by filtering or joining, as in the traditional relational algebra, but also by the imputation operators.
For example, a drop operator will reduce the cardinality of the result while an impute operator will maintain the same cardinality as the input.

@@

\todo{Add a verbal explanation of the scaling steps}
\begin{algorithm}
\input{algorithms/histogram-transform.tex}
\end{algorithm}
